---
title: "FutureTenseConstructionL2English-DataProcessing"
author: "Daniel Crawford"
date: "04/02/2024"
output: 
  github_document: 
    toc: TRUE
---


# Correlation of Future Tense Construction Preference with Proficiency Scores for English L2 Learners
```{r, setup}
knitr::opts_chunk$set(fig.path = paste0(dirname(getwd()),'/images/DataAnalysis-'))
```

```{r, label = 'import'}
#Import Packages
library(tidyverse)
```

## Load Data
Raw data can be found [here](https://github.com/ELI-Data-Mining-Group/PELIC-dataset/raw/master/PELIC_compiled.csv).

### Read in Data from PELIC
```{r, label = 'readPelic'}
#Read in Data from PELIC: 

if (file.exists(paste0(dirname(getwd()),'/large_data/pelic_compiled.csv'))){
  print('Reading in File from directory')
  PELIC_compiled = read.csv(paste0(dirname(getwd()),'/large_data/pelic_compiled.csv'))
}else{
  print('Reading in File from url')
  PELIC_compiled = as_tibble(read.csv(url("https://github.com/ELI-Data-Mining-Group/PELIC-dataset/raw/master/PELIC_compiled.csv"), fileEncoding = "ISO-8859-1"))
}

nrow(PELIC_compiled)
ncol(PELIC_compiled)
head(PELIC_compiled)
```

The data contains about 46,000 entries, and a vast about of background data on the students. Of note will be L1, level, and scores on proficiency measures. 


### Read in Student Indo
```{r, label = 'readStudentInfo'}
#Need to get demographic and proficiency scores for each student

student_info = as_tibble(read.csv(url("https://github.com/ELI-Data-Mining-Group/PELIC-dataset/raw/master/corpus_files/student_information.csv")))
nrow(student_info)
ncol(student_info)
head(student_info)
  
```

The student information for the 1,313 students expands on the demographics, and will be joined in later to the data. 


## Tokenize Data

### Tokenize Lemmatized Data Helper Function
```{r, label = 'tokFunc'}
#A function to create a tokenized df with columns: token | lemma | POS
#Input is a string
tlp_text_to_df = function(x){
  text_df = x %>% 
    #Remove "[" character (at beginning of string)
    str_remove("^\\[\\(") %>% 
    #Remove "]" character (at end of string)
    str_remove("\\)\\]$") %>% 
    #split tuples
    str_split_1("\\), \\(") %>% 
    #remove 's
    str_remove_all("'") %>% 
    #convert to tibble
    as.tibble() %>% 
    #remove rows of commas - these are not relevant or impactful on analysis
    filter(!startsWith(value, ",,")) %>% 
    #separate into columns
    separate(value, into = c('token','lemma','POS'), sep = ',')
  
  return(text_df)
  
}
```


```{r, label = 'readTokUnnested'}
if (file.exists(paste0(dirname(getwd()),'/large_data/tokenized_df_unnested.csv'))){
  print('Reading in File from directory')
  tokenized_df_unnested = read.csv(paste0(dirname(getwd()),'/large_data/tokenized_df_unnested.csv'))
}else{
  
  print('Creating tokenized_df_unnested - about 10 minute runtime.')
  # !!! WARNING - this will take about 10 - 15 !!!
  
  #Create tokenized_df, each row is a token
  tokenized_df_unnested = PELIC_compiled %>%
   #create tokenized df from the tok_lem_POS string
   mutate(tokenized_nested_df = map(tok_lem_POS, tlp_text_to_df)) %>%
   #keep the unique identifier of answer id
   select(answer_id, tokenized_nested_df) %>%
   #make the df longer and unndest
   unnest(tokenized_nested_df)

  # un-comment if you want to save
  #tokenized_df_unnested %>% write.csv('large_data/tokenized_df_unnested.csv')
}
```

## Summary of Data 

### Summary of 'will' tokens
```{r, label = 'countWill'}
tokenized_df_unnested %>% 
  filter(trimws(lemma) == "will" & trimws(token) != 'will' ) %>% 
  count(token, lemma)
```

### Summary of 'going' tokens
```{r, label = 'summaryGoing'}
tokenized_df_unnested %>% 
  filter(trimws(lemma) == "go" & trimws(token) != 'going' ) %>% 
  count(token)
```
Note that not all of these are of the 'going to' future construction

## Format the Data 
```{r, label = 'formatData'}

#Create counts_data, each row is a token
counts_data =  tokenized_df_unnested %>% 
  #if there is a 'will' token that is a modal, put 1 in new column, else 0
  mutate(is_will_construction = if_else(trimws(lemma) == 'will' & trimws(POS) == 'MD', 1,0)) %>% 
  #if there is a 'going' token followed by 'to' then a verb within 2 words, put 1 in new column
  mutate(is_goingTo_construction = 
            if_else(
              token == "going" & lead(token) == "to" & (
                #lead allows to 'look ahead'
                startsWith(trimws(lead(POS, 2)), 'V') | 
                startsWith(trimws(lead(POS, 3)), 'V')
              ),1,0)) %>% 
  #get the answer id and counts
  select(answer_id, is_will_construction, is_goingTo_construction) %>%
  group_by(answer_id) %>% 
  #sum counts by answer ID
  summarise(
    count_will_construction = sum(is_will_construction), 
    count_goingTo_construction = sum(is_goingTo_construction)
  )

#this is the final data frame with the students info
countruction_counts_and_student_info  = PELIC_compiled %>% 
  #joing in the tokenized df with counts, on answer ID
  left_join(counts_data, by = "answer_id") %>% 
  #get the anon_id and counts, we do not need the text info
  select(anon_id, count_will_construction, count_goingTo_construction) %>% 
  group_by(anon_id) %>% 
  #count construction usage by student
  summarise(
    count_will_construction = sum(count_will_construction), 
    count_goingTo_construction = sum(count_goingTo_construction)
  ) %>% 
  #bring in student info
  left_join(student_info, by = "anon_id")
  

        
```

### Bring in score data

```{r, label = 'joinScores'}
c

#Find students who took the test more than once
n_occur <- data.frame(table(score_data$anon_id))
multiple_ids = n_occur[n_occur$Freq > 1,]$Var1

#join in scores of students
countruction_counts_and_student_info_with_scores = countruction_counts_and_student_info %>% 
  inner_join(score_data %>% filter(!(trimws(anon_id) %in% multiple_ids)), by = 'anon_id')


```




### Save The data

```{r, label = 'saveConstrCounts'}
countruction_counts_and_student_info_with_scores %>% 
  write.csv(paste0(dirname(getwd()),'/Data Files/FINAL_DATA_countruction_counts_and_student_info_with_scores.csv'))
```


My strategy for importing the data is to find construction in question in the text, and then to join in the student demographic information


```{r, label = 'saveTokenizedData'}
#In order to perform analysis on tokens later
filtered_tokenized_data = tokenized_df_unnested %>% 
  #if there is a 'will' token that is a modal, put 1 in new column, else 0
  mutate(is_will_construction = if_else(trimws(lemma) == 'will' & trimws(POS) == 'MD', 1,0)) %>% 
  #if there is a 'going' token followed by 'to' then a verb within 2 words, put 1 in new column
  mutate(is_goingTo_construction = 
            if_else(
              token == "going" & lead(token) == "to" & (
                #lead allows to 'look ahead'
                startsWith(trimws(lead(POS, 2)), 'V') | 
                startsWith(trimws(lead(POS, 3)), 'V')
              ),1,0)) %>% 
  filter(is_will_construction | is_goingTo_construction)

filtered_tokenized_data %>% write.csv(paste0(dirname(getwd()),'/Data Files/filtered_tokenized_data.csv'))
```


```{r, label = 'countShall'}
tokenized_df_unnested %>% 
  filter(trimws(lemma) == 'shall')
```




## Session Info
```{r}
sessionInfo()
```

